---
title: "Lab 13: Large Sample CI"
output:
  pdf_document:
    keep_tex: true
header-includes:
   - \usepackage{booktabs}
geometry: margin=1.5cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spatial Organization of Chromosome (Rice Problem 8.45)

The R code below reads in the data and calculates the maximum likelihood estimate:

```{r}
library(readr)
library(ggplot2)

chromatin <- read_csv("http://www.evanlray.com/data/rice/Chapter%208/Chromatin/data05.txt",
  col_names = FALSE)
colnames(chromatin) <- "distance"

theta_hat <- 1/(2 * nrow(chromatin)) * sum(chromatin$distance^2)
theta_hat
```

#### 1. Find the variance of a large-sample normal approximation to the sampling distribution of the MLE based on the observed Fisher information.  (You should be able to calculate a number.)

```{r}
est_var_theta_hat_J <- 1 / (1/theta_hat^3 * sum(chromatin$distance^2) - nrow(chromatin) / theta_hat^2)
est_var_theta_hat_J
```

#### 2. Find the variance of a large-sample normal approximation to the sampling distribution of the MLE based on the Fisher information.  (You should be able to calculate a number.)

```{r}
est_var_theta_hat_I <- theta_hat^2 / nrow(chromatin)
est_var_theta_hat_I
```

#### 3. The plot below shows a representation of the pdf of the $\text{Rayleigh}(\theta)$ distribution based on the maximum likelihood estimate of $\theta$.  Add two more curves corresponding to the values of $\theta$ at the upper and lower end points of an approximate 95% confidence interval for $\theta$ based on the Fisher information from part 2.

R does not come with a function to calculate the pdf of the Rayleigh distribution, so I have defined one below for you to use.

```{r, fig.height = 3}
drayleigh <- function(x, theta) {
  log_value <- rep(-Inf, length(x))
  inds <- which(x > 0)
  
  log_value[inds] <- log(x[inds]) - log(theta) - (x[inds]^2)/(2 * theta)
  
  return(exp(log_value))
}




ggplot(data = chromatin, mapping = aes(x = distance)) +
  geom_histogram(binwidth = 0.5, mapping = aes(y = ..density..)) +
  stat_function(fun = drayleigh, args = list(theta = theta_hat), color = "orange") +
  stat_function(fun = drayleigh, args = list(theta = theta_hat - qnorm(0.025) * sqrt(est_var_theta_hat_I)), color = "cornflowerblue") +
  stat_function(fun = drayleigh, args = list(theta = theta_hat - qnorm(0.975) * sqrt(est_var_theta_hat_I)), color = "purple")
```
